Parsing Postfix log files

### 2009-04-08 doc/logparser-results.tex
Finished (I hope) the minimum vs maximum ruleset section.

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full \parsername{} ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
used when generating the results in this thesis is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because \parsername{} is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses log files
generated by later Postfix versions.  A second set of test runs was
performed using the minimum ruleset, and the parsing times compared to
those generated using the full ruleset: the percentage parsing time
increase when using the full ruleset instead of the minimal ruleset for
optimal, shuffled and reverse orderings is shown in \graphref{Percentage
parsing time increase of maximum ruleset over minimum ruleset}.

@@

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering.  The
optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
XXXXXXXXXXXX IMPROVE THIS EXPLANATION\@.  Log files 22 and 62--68 show much
smaller increases for shuffled and reversed ordering than other log files
do, because the majority of log lines in those log files are parsed by
Postfix components with a small number of parsing rules, so removing
unnecessary rules has little effect on the total number of rules used.
These results show that the architecture scales very well as the number of
rules increases, and that optimally sorting the rules is an important
optimisation contributing to this scalability.

@@

XXX UPDATE THIS\@.  Two large dips can be seen in \graphref{Increase in
parsing time when not caching compiled regexes graph} at log files 22 and
62--68, corresponding to the spikes in log file size in \graphref{parsing
time vs file size vs number of log lines graph}.  The distribution of log
lines across rules when there is a mail loop is much different to the norm,
and the average number of rules consulted per log line is much lower: this
results in far fewer regex compilations per line when there is a mail loop,
and a corresponding decrease in the average parsing time for a single log
line.  The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

@@

>>>>

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full \parsername{} ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
used when generating the data in this chapter is \numberOFrulesMINIMUM{}
rules, \numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full
ruleset is larger because \parsername{} is tested with
\numberOFlogFILESall{} log files; testing with more log files increases the
chance of finding bugs in the parser or new complications to be overcome.
The \numberOFlogFILES{} log files were each parsed 10 times using the
minimum ruleset, and the parsing times compared to those generated using
the full ruleset: the percentage parsing time increase when using the full
ruleset instead of the minimal ruleset for optimal, shuffled and reverse
orderings is shown in \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset}; mean and standard deviation are shown in
\tableref{Percentage parsing time increase of maximum ruleset over minimum
ruleset table}.

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering, whereas
optimal ordering shows very little change.  Log files 22 \& 62--68 show
much smaller increases than other log files do, because the majority of log
lines in those log files are produced by Postfix components with a small
number of rules, so removing unnecessary rules has little effect on the
total number of rules used; \tableref{Number of rules per Postfix component
in the maximum and minimum rulesets} shows the number of rules per Postfix
component for each ruleset.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that the architecture scales very well as the number of
rules increases, and that optimally sorting the rules is an important
optimisation contributing to this scalability.

@@

\showtable{build/include-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset table}

\begin{table}[thbp]
    \caption{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \empty{}\label{Number of rules per Postfix component in the maximum and
    minimum rulesets}
    \centering{}
    \begin{tabular}{lrr}
        \tabletopline{}%
        Postfix component & Maximum ruleset & Minimum ruleset \\
        \tablemiddleline{}%
        \input{build/include-number-of-rules-per-program-minimum-ruleset}
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

XXX UPDATE THIS\@.  Two large dips can be seen in \graphref{Increase in
parsing time when not caching compiled regexes graph} at log files 22 \&
62--68, corresponding to the spikes in log file size in \graphref{parsing
time vs file size vs number of log lines graph}.  The distribution of log
lines across rules when there is a mail loop is much different to the norm,
and the average number of rules consulted per log line is much lower: this
results in far fewer regex compilations per line when there is a mail loop,
and a corresponding decrease in the average parsing time for a single log
line.  The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

@@

XXX COMMENT THAT THE OPTIMISATION TIME HAS BEEN SPENT ON RULES, AND THAT IS
GOOD BECAUSE THAT IS WHERE THE MAJORITY OF EXECUTION TIME IS SPENT\@.
ALSO, RULE OPTIMISATION IS GENERIC, WHEREAS ACTION OPTIMISATION IS
PARSER-SPECIFIC\@.

### 2009-04-08 doc/logparser.tex
Updated results after stats-max run finished. Added perfect best and perfect worst graph, and a table showing the std dev etc. Worked on graph and table placement, without much success.  I think manual     tweaking will be required, though adding 'h' back into the placement list     has helped. Added a graph for the number of times each action is invoked.
Added a table showing mean and stddev for the number of invocations of each action, for different groups of log files.
Use the booktabs packages to provide better horizontal rules for tables.
Removed some make todo false positives. Removed so

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

@@

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Warnings rather than errors when references cross page boundaries.
\vrefwarning{}
% Errors when loops are encountered
\vrefshowerrors{}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

@@

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the defaul spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    % XXX REVISIT THE PLACEMENT OF FIGURES AND TABLES, THOUGH I THINK
    % LATEX'S DEFAULTS ARE PROBABLY PRETTY GOOD.
    \begin{figure}%[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[3]{
    \begin{table}%[ht]
        \caption{#2}\label{#3}
        \input{#1}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \hline%
    \noalign{\smallskip}%
}

\newcommand{\tablebottomline}[0]{%
    \noalign{\smallskip}%
    \hline%
}

\newcommand{\tablemiddleline}[0]{%
    \noalign{\smallskip}%
    \hline%
    \noalign{\smallskip}%
}

@@

% XXX UPDATE THIS WHEN all-logs PASSES
\newcommand{\numberOFlogFILESall}[0]{%
    522%
}

@@

    XXX FIX THE PROBLEMS THAT AROSE FROM THE LAST all-logs RUN\@.

    XXX RUN ``test-some'' AGAIN TO UPDATE THE DB\@.

    XXX SHOULD I HAVE A ``FUTURE WORK'' SECTION SOMEWHERE IN THE
    CONCLUSION\@?

@@

>>>>

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Provides \FloatBarrier that stops floats going past it.
\usepackage{placeins}
% Relax the parameters Latex uses when placing floats.
\usepackage{flexiblefloats}
% Prevent floats from appearing above the line that includes them
\usepackage{flafter}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Better looking horizontal rules for tables.
\usepackage{booktabs}

@@

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

@@

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the default spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
    \setlength{\topsep}{0pt}%
    \setlength{\partopsep}{0pt}%
}

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[4][thbp]{%
    \begin{figure}[#1]
        \caption{#3}\label{#4}
        \includegraphics{#2}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[4][thbp]{%
    \begin{table}[#1]
        \caption{#3}\label{#4}
        \input{#2}
    \end{table}
}

\newcommand{\tabletopline}[0]{%
    \toprule{}%
}

\newcommand{\tablebottomline}[0]{%
    \bottomrule{}%
}

\newcommand{\tablemiddleline}[0]{%
    \midrule{}%
}

@@

\newcommand{\numberOFlogFILESall}[0]{%
    774%
}

@@

    XXX ASK CARL\@: SHOULD I HAVE A ``FUTURE WORK'' SECTION SOMEWHERE IN
    THE CONCLUSION\@?

@@

    XXX SPELL CHECK AND RUN DICTION\@.

    XXX MAKE IT EASY TO REDUCE THE HEIGHT OF GRAPHS\@.
